---
title: "p8105_hw3_th3072"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

```

# Problem 1

## part 1
```{r}
data("instacart")
# Number of aisles
number_aisles <- nrow(instacart)

# Aisles with most orders
aisles_most_ordered <- instacart |>
  group_by(aisle) |>
  summarize(count = n()) |>
  arrange(-count)
```

*Answer*

There are `r nrow(instacart)` aisles and fresh vegetables are the most items ordered from.

## part 2
```{r}
filtered_aisles <- aisles_most_ordered |>
  filter(count > 10000)

ggplot(filtered_aisles, aes(x = reorder(aisle, -count), y = count)) +
  geom_point() +
  coord_flip() +
  labs(title = "Number of Items Ordered by Aisle", x = "Aisle", y = "Number of Orders")
```
## part 3
```{r}
# Three most popular items in each of the aisle
popular_items <- instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |>
  summarize(count = n(), .groups = 'drop') |>
  arrange(aisle, -count) |>
  group_by(aisle) |>
  slice_head(n = 3)

```

```{r}
# mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
mean_hour <- instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = product_name,
    values_from = mean_hour
  )
```

# Problem 2

```{r}
# Clean the data
data("brfss_smart2010")
brfss_cleaned <- brfss_smart2010 |>
  select(Year, Locationabbr, Locationdesc, Topic, Response, Data_value) |>
  filter(Topic == "Overall Health", 
         Response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) |>
  mutate(Response = factor(Response, 
                           levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"), 
                           ordered = TRUE))

```

## part 1
```{r}
# state were observed at 7 or more location (2002 and 2010)
seven_or_more_observed <- brfss_cleaned |>
  group_by(Year, Locationabbr)|>
  summarize(location = n_distinct(Locationdesc))|>
  filter(location >=7 & (Year == 2002 | Year == 2010))
```

*Answer*

In 2002, CT, FL, MA, NC, NJ and PA were observed at 7 or more locations.

In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA were observed at 7 or more locations.

## part 2
```{r}
excellent_data <- brfss_cleaned |>
  filter(Response == "Excellent") |>
  group_by(Year, Locationabbr) |>
  summarize(mean_data_value = mean(Data_value, na.rm = TRUE))

ggplot(excellent_data, aes(x = Year, y = mean_data_value, group = Locationabbr, color = Locationabbr)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Average Data Value for 'Excellent' Responses Over Time", 
       y = "Average Data Value")

```
## part 3
```{r}
ny_data <- brfss_cleaned |>
  filter(Locationabbr == "NY" & (Year == 2006 | Year == 2010))

ggplot(ny_data, aes(x = Response, y = Data_value)) +
  geom_boxplot() +
  facet_grid(. ~ Year) +
  theme_minimal() +
  labs(title = "Distribution of Data Value for Responses in NY State (2006 & 2010)", 
       y = "Data Value")

```



# Problem 3
```{r}
demo_data <- read.csv("./hw3_data/nhanes_covar.csv", skip = 4)
accel_data <- read.csv("./hw3_data/nhanes_accel.csv")

# Merge the two datasets 
demo_accel_data <- merge(demo_data, accel_data, by = "SEQN")

```

```{r}
demo_accel_data$sex <- ifelse(demo_accel_data$sex ==1, "male", "female")
demo_accel_data$education <- ifelse(demo_accel_data$education ==1, "Less than high school",
                            ifelse(demo_accel_data$education == 2, "High school equivalent", "More than high school"))
cleaned_data <- demo_accel_data |>
  filter(age >= 21, !is.na(sex), !is.na(education), !is.na(BMI), !is.na(age)) |>
  mutate(sex = factor(sex, levels = c("male", "female")),
         education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school")))
```

```{r}
education_gender <- cleaned_data |>
  group_by(education, sex) |>
  summarize(count = n())

```

## part 1
```{r}
# number of men and women in each education category and create a visulization of the age distributions for men and women in each education category
cleaned_data |>
  ggplot(aes(x = age, color = sex)) +
  geom_histogram(binwidth = 5) +
  facet_grid(~ education) +
  labs(title = "Age distributions by sex and education level",
       x = "Age",
       y = "Count") +
  theme_minimal()

```
*Answer*

In this table, we can clearly see that in this dataset, the number of people who is `more than high school` is more than `High school equivalent` and `Less than high school`. Moreover, regardless of age, there are more male than females in this dataset.

## part 2
```{r}
# Aggregate activity data
activity_summary <- cleaned_data|>
  group_by(SEQN,education,sex,age)|>
  pivot_longer(
  min1:min1440,
  names_to = "time",
  values_to = "activity"
  )|>
  summarize(totalactivity = sum (activity))

# Plot total activities against age
ggplot(activity_summary, aes(x = age, y = totalactivity, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(~education) +
  labs(x = "Age", y = "Total Activity", color = "sex") +
  theme_minimal()
```

*Answer*

Regardless of the educational category, Total Activity decreases as the age increase. 

```{r}
twentyfour_hour_activity <- cleaned_data|>
  group_by(SEQN,education, sex)|>
  pivot_longer(
  min1:min1440,
  names_to = "time",
  values_to = "activity"
  )

# Plot 24-hour activity time courses
ggplot(twentyfour_hour_activity, aes(x = time, y = activity, color = sex)) +
  geom_smooth(method = "lm") +
  geom_point()+
  facet_wrap(~education) +
  labs(title = "24-hour activity", x = "time", y = "activity", color = "sex") +
  theme_minimal()

```
*Answer*

Regardless of the educational category, 24 hour activity is almost the same. 
